Meta-analysis

### 2020-07-23 04_MetaAnalysis.Rmd
Make corrections and fix so it knits

   - Grain yield is our response variable and will be converted to kg / ha

@@

   - We will test spray management (our treatment) as a fixed effect and random slope to trial

@@

kable(
   table(slimmer_PM_dat$spray_management, slimmer_PM_dat$year),
   align = rep('c', 8),
   caption = "Which treatments and how many treatments are represented in each year"
) %>%
   kable_styling(
      "striped",
      fixed_thead = TRUE,
      full_width = FALSE,
      position = "center"
   )
```

@@

There seems like little difference between the treatments, with exception of `Late_plus`.
Let's do this plot again, but let's look at the proportional mean difference between the treatments and the no spray control for each study, this should reduce variation in yield due between trials.

@@

Using the mean difference in the treatment effect seems to show the effect of each treatment better than just raw yield and reduces the variability in the treatments.
We know there is a good deal of variability in our studies and between them.
Mungbean produces variable yields between seasons so we should use a response that highlights the difference in the treatment effects that we are interested in and reduce the variability.
Let's calculate the standardised mean difference for each treatment to reduce the variability in the meta-analysis.

@@

We have removed the studies that don't report any variance from the analysis.

### Calculate standardised mean differences

Let's calculate standardised mean differences and add them to the data frame.

```{r StandardisedMeanDifference}
slimmer_PM_dat$pooledSD <- NA
slimmer_PM_dat$vi_C <- NA
for (i in slimmer_PM_dat$trial_ref) {
   # First we need to Calculate the pooled standard deviation
   slimmer_PM_dat[slimmer_PM_dat$trial_ref == i, "pooledSD"] <-
      sqrt(sum((slimmer_PM_dat[slimmer_PM_dat$trial_ref == i, "n"] - 1) * slimmer_PM_dat[slimmer_PM_dat$trial_ref == i, "yield_error"]) /
              (sum(slimmer_PM_dat[slimmer_PM_dat$trial_ref == i, "n"]) - nrow(slimmer_PM_dat[slimmer_PM_dat$trial_ref == i, ])))

   # Create a column with the value of the mean grain yield of the no spray
   #  control for comparisons to the treatments means
   slimmer_PM_dat[slimmer_PM_dat$trial_ref == i, "yi_C"] <-
      mean(slimmer_PM_dat[slimmer_PM_dat$trial_ref == i &
                             slimmer_PM_dat$fungicide_ai == "control",
                          "grain_yield.t.ha"], na.rm = TRUE)

}

# Calculate standardised mean difference
slimmer_PM_dat$grain_SMD <-
   (slimmer_PM_dat$grain_yield.t.ha - slimmer_PM_dat$yi_C) / slimmer_PM_dat$pooledSD
```

To visualise this formula in mathematical notation:  
$SMD = \frac{Y_i-\overline{Y_c}}{\sigma_p}$

We can use the same sample variance with the standardised mean difference.
This is because if you add or subtract from a random variable, the variance does not change.

Let's simplify our data by removing the control data, converting disease pressure (`D_pres`) to a factor and selecting only the columns of data necessary for analysis.
We will also include the mean difference or response ratio `grain_MD`, between the treatment and control as a comparison of responses.

```{r metafor_organisation}
dat1 <-
   slimmer_PM_dat %>%
   filter(fungicide_ai != "control") %>% #remove controls from the data
   mutate(spray_management =
             factor(
                spray_management,
                c(
                   "Early",
                   "Recommended",
                   "Recommended_plus",
                   "Late",
                   "Late_plus"
                )
             ),
          D_pres = 
             factor(D_pres,
                    c("lowD",
                      "highD")),
          grain_MD = grain_yield.t.ha - yi_C) %>%
   select(
      trial,
      trial_ref,
      location,
      year,
      row_spacing,
      host_genotype ,
      spray_management,
      fungicide_ai,
      D_pres,
      grain_yield.t.ha,
      grain_SMD,
      grain_MD,
      dose,
      vi,
      vi2
      )
```

@@

Let's inspect the data to determine if we need to transform the response variable.

```{r}
hist_usq(dat1$grain_yield.t.ha)
hist_usq(log(dat1$grain_yield.t.ha))
hist_usq(sqrt(dat1$grain_yield.t.ha))
hist_usq(dat1$grain_MD)
```

Standardised mean differences have a bit of a long tail however this is not bad and transformations with log or square root will not work on negative values.
So it looks like no transformation is necessary before we start the meta-analysis.

Let's undertake the meta-analysis using the package `metafor`.
We are using the `spray_management` variable as a moderator and an interactive term to the `trial` random variable.
Because we want to know the difference between all treatments and we have no reference treatment, we will remove the intercept.

```{r Metafor-analysis}
PM_mv <- rma.mv(
   yi = grain_MD,
   vi,
   mods = ~ spray_management - 1,
   method = "ML",
   random = ~ spray_management | trial,
   struct = "UN",
   data = dat1
)
summary(PM_mv)
```

Let's see if we can improve this model by including some other random effects and improving the random effect structure in the model.

```{r rma_noRandomInteractiveTerm}
PM_mv_St <- rma.mv(
   yi = grain_MD,
   vi,
   mods = ~ spray_management,
   method = "ML",
   random = list( ~ 1 | trial),
   data = dat1
)

PM_mv_St1 <- rma.mv(
   yi = grain_MD,
   vi,
   mods = ~ spray_management,
   method = "ML",
   random = list(
      ~ 1 |
         location / year / trial_ref / dose / host_genotype / factor(row_spacing) /
         spray_management
   ),
   data = dat1
)

AIC(PM_mv) # Printing the AIC of the model helps differentiate the models in the anova output
anova(PM_mv, PM_mv_St) # PM_mv is the better model
AIC(PM_mv_St)
anova(PM_mv_St, PM_mv_St1) # PM_mv_St is the better model
```

The simpler `PM_mv_St` model explains significantly less variation than the more complicated model `PM_mv`.
So we will keep the original model `PM_mv`.

I have also tested the structure for testing the random trial and spray_management variables, these show that there is no significant difference to the models `PM_mv_St` and `PM_mv_St1` and that the originally method of specifying the random variables within the `trial` factor provides us with the best model.

@@

@@

The fixed effects, in the last table, show that yields in single early spray treatments are not significantly different from zero, or the no spray control.

@@

*****

### Metafor comparisons

To make it easier to compare each of the treatments we can compute the meta-analysis contrasts.

```{r metafor_contrasts}
meta_cont <- anova(PM_mv, L = rbind(
   c(-1, 1, 0, 0, 0),
   c(-1, 0, 1, 0, 0),
   c(-1, 0, 0, 1, 0),
   c(-1, 0, 0, 0, 1),
   c(0, -1, 1, 0, 0),
   c(0, -1, 0, 1, 0),
   c(0, -1, 0, 0, 1),
   c(0, 0, -1, 1, 0),
   c(0, 0, -1, 0, 1),
   c(0, 0, 0, -1, 1)
))

change_labels <- function(x1) {
   x1 <- gsub(pattern = "spray_management",
              replacement = "",
              x = x1)
   x1
}

source("R/p_star.R")

table3 <- data.frame(
   contrasts = change_labels(meta_cont$hyp[, 1]),
   estimates = meta_cont$Lb,
   se = meta_cont$se,
   z_val = meta_cont$zval,
   pval = meta_cont$pval,
   VisP = p_star(meta_cont$pval)
)
write.csv(table3, here("cache/Table3_contrasts.csv"), row.names = FALSE)
```

Results show with the exclusion of the no spray `control` and `Late` treatments, `Early` applications are significantly worse than all other treatments leading to lower yield estimates.
`Late_plus` is estimated to save the most yield however it is only significantly higher than `Early` and `Late` treatments.
Let's put these values in a plot to visualise the differences between treatments.

First we will format the results into a data frame.

@@

```{r metafor_plot}
efficacy %>%
   mutate(Treatment = factor(Treatment, levels = rev(
      c(
         "Early",
         "Recommended",
         "Recommended_plus",
         "Late",
         "Late_plus"
      )
   ))) %>%
   ggplot(aes(Treatment, Mean)) +
   geom_hline(
      yintercept = seq(-0.05, 0.3, by = 0.05),
      color = usq_cols("usq charcoal"),
      linetype = 3
   ) +
   geom_hline(yintercept = 0) +
   geom_point(aes(size = 1 / SE), shape = 15) +
   geom_linerange(aes(ymin = CIs_lower, ymax = CI_upper)) +
   coord_flip()
```

### Disease pressure effect

We also want to know if disease pressure exerts an effect on the decision for spray control.
Above we can see spraying at the first sign of powdery mildew or with in `r min(as.numeric(slimmer_PM_dat[slimmer_PM_dat$spray_management == "Late"| slimmer_PM_dat$spray_management == "Late_plus","fungicide_timing_1"]))` - `r max(as.numeric(slimmer_PM_dat[slimmer_PM_dat$spray_management == "Late"|                   slimmer_PM_dat$spray_management == "Late_plus","fungicide_timing_1"]))` days after first sign is effective at mitigating yield loss.
Does this change between years of low and high disease pressure.

We will add the intercept back into this model for ease of interpretation. 
```{r D_pres-Meta}
PM_mv_Dp <- rma.mv(
   yi = grain_MD,
   vi,
   mods = ~ spray_management * D_pres - 1,
   method = "ML",
   random = ~ spray_management | trial,
   struct = "UN",
   data = dat1
)

summary(PM_mv_Dp)
```

These results show that the spray management variable we tested are not significantly better or worse between the two levels of disease pressure.
It seems that disease preasure does not improve the models ability to assess the most effective fungicide spray management option. Let's compare the disease pressure model to the original model to test this assumption.

```{r compareDvNDP}
AIC(PM_mv_Dp)
anova(PM_mv_Dp, btt = "D_pres")
anova(PM_mv, PM_mv_Dp)
```

A Wald test of `D_pres` shows the moderator is not significant in accounting for heterogeneity (P = 0.4859).
No significant difference which informs us to use the original model `PM_mv`, which is the simplest.

### Test for bias in indirect comparisons 
Treatments `Early` and `LatePlus` did not co-occur in any of the trials considered in the meta-analysis.
Indirect comparisons are ok to interpret in the results of a network meta-analysis, however if there are biases between trials including `Early` or `LatePlus` treatments then the results are likely bias. 
Therefore we wish to do a quick check that the trials which included `Early` treatments did not significantly differ from trials which incorportated `LatePlus`,
Lets create a data.frame of values for this comparison

```{r}
Ear <- 
   dat1 %>%
   filter(spray_management == "Early")%>%
   distinct(trial_ref)

LP <- 
   dat1 %>%
   filter(spray_management == "Late_plus")%>%
   distinct(trial_ref)

dat1$indirectC <- NA

dat1 %>%
   filter(trial_ref == Ear$trial_ref)

dat1[dat1$trial_ref == Ear$trial_ref, "indirectC"] <- "Ear"
dat1[dat1$trial_ref == LP$trial_ref, "indirectC"] <- "LP"

dat1 %>%
   filter(!is.na(indirectC))%>%
   ggplot(aes(x = indirectC, y = grain_MD))+
   geom_boxplot()

dat1 %>%
   filter(!is.na(indirectC),
          spray_management != "Early",
          spray_management != "Late_plus")%>%
   ggplot(aes(x = indirectC, y = grain_MD))+
   geom_boxplot()

dat1[dat1$trial_ref == LP$trial_ref, "indirectC"] <- "LP"
dat1[dat1$trial_ref == Ear$trial_ref, "indirectC"] <- "Ear"

dat2 <- dat1 %>%
   filter(!is.na(indirectC),
          spray_management != "Early",
          spray_management != "Late_plus")

table(dat2$spray_management, dat2$indirectC)

summary(aov(grain_MD ~ indirectC + spray_management + trial_ref, data = dat2))

@@

This anova shows that there is no significant difference between the trials, spray_management or the specific trials with either `Early` or `Late_plus`. 
Therefore there is low risk that interpreting results from this indirect comparison between `Early` and `Late_plus`.

@@

```{r Save_meta_data, eval=FALSE}
write.csv(slimmer_PM_dat, file = "data/GYmeta_data.csv")
```

>>>>

   - Grain yield is our response variable (t / ha)

@@

   - We will test spray management (our treatment) as a fixed effect and random interactive term to trial

@@

kable(
   table(slimmer_PM_dat$spray_management, slimmer_PM_dat$year),
   align = rep('c', 8),
   caption = "Which treatments and how many treatments are represented in each year"
) %>%
   kable_styling(
      "striped",
      fixed_thead = TRUE,
      full_width = FALSE,
      position = "center"
   )

```
<!-- I don't like this graph, it is hard to read given that different widths don't translate to anything meaningful -->

@@

There seems like little difference between the treatments, with exception of `Late_plus`.
Let's do this plot again, but let's look at the proportional mean difference between each treatment and the no spray control for each study, this should reduce variation in yield due between trials.

@@

Using the mean difference in the treatment effect seems to show the effect of each treatment better than just raw yield and reduces the variability in the treatments.
We know there is a good deal of variability in our studies and between them.
Mungbean produces variable yields between seasons so ordinarilly we should use a response that highlights the difference in the treatment effects that we are interested in and reduce the variability.
However for an arm-based analysis we calculate the differences between control and treatment after the meta-analysis.

### Formating the response variable

Let's inspect the response variance `grain_yield.t.ha` to see if it is normally distributed. 
In many cases a log transformation of the response variable improves the fit, due to the prevalence of log distributed data in nature.

```{r}
hist_usq(slimmer_PM_dat$grain_yield.t.ha, main = "Grain yield (t / ha)")
hist_usq(log(slimmer_PM_dat$grain_yield.t.ha), main = "log(Grain yield (t / ha))")
hist_usq(sqrt(slimmer_PM_dat$grain_yield.t.ha), main = "sqrt(Grain yield (t / ha))")
hist_usq(slimmer_PM_dat$yield_gain, main = "Mean difference to control (t / ha)")
```
The distribution of data does not seem to be improved by transformation of the data.
We will retain the continuous variable grain_yield.t.ha as the response.
Mean differences won't be used as we are following the methodology of a ARM-based model, where mean differences are calculated following the meta-analysis.

@@

Now Let's simplify our data by removing the control data, converting disease pressure (`D_pres`) to a factor and selecting only the columns of data necessary for analysis.
We will also include the mean difference or response ratio `grain_MD`, between the treatment and control as a comparison of responses.

```{r metafor_organisation}
dat1 <-
   slimmer_PM_dat %>%
   mutate(vi = vi/n) %>% 
   mutate(spray_management =
             factor(
                spray_management,
                c(
                   "control",
                   "Early",
                   "Recommended",
                   "Recommended_plus",
                   "Late",
                   "Late_plus"
                )
             ),
          D_pres = 
             factor(D_pres,
                    c("lowD",
                      "highD")),
          id = row_number()) %>%
   dplyr::select(
      trial,
      trial_ref,
      location,
      year,
      row_spacing,
      host_genotype,
      spray_management,
      fungicide_ai,
      D_pres,
      grain_yield.t.ha,
      yield_gain,
      dose,
      vi,
      yield_error,
      vi2,
      n,
      id)
```

@@

To do a ARM based model we will use the package `metafor`.
We are using the `spray_management` variable as a moderator and an interactive term to the `trial` random variable.

```{r Metafor-analysis}
PM_mv <- rma.mv(
   yi = grain_yield.t.ha,
   vi,
   mods = ~ spray_management,
   method = "ML",
   random = list(~ spray_management | trial, ~1 | id), 
   struct = "UN",
   control=list(optimizer="optim"),
   data = dat1
)

summary(PM_mv)
```

@@

The tau^2 also gives us the heterogeneity between trials and indicates `Recommended_plus` showed the highest heterogeneity between experiments followed by `Late_plus`, `Late` then `Recommended`, with `Early` and the no-spray `control` showing the least heterogeneity.

@@

In this result we can see that the `Early` treatment is not significantly different to the intercept, which in this case is the mean of the no spray `control`.
However the other treatments are significantly different from the no-spray control (intercept).
The $Q_M$ [omnibus test](http://www.metafor-project.org/doku.php/tips:models_with_or_without_intercept?s[]=anova) of moderators, shows the moderators significantly influence the model ($Q_M =$ `r PM_mv$QM` $, df = $ `r PM_mv$m`, `r reportP(PM_mv$QMp)`) and we can reject the null hypothesis ($H_0 : \beta_1 = \beta_2 = \beta_3 =\beta_4 = 0$) that there is no difference between the moderators [@Viechtbauer2010].
The analysis shows there is still a significant amount of residual heterogeneity ($Q_E = $`r PM_mv$QM` $, df = $`r PM_mv$k - PM_mv$m`, `r reportP(PM_mv$QEp)` ) not captured by the spray management moderator indicating other possible moderators which might influence grain yield.

```{r PM_Mv_Contrast}
source("R/simple_summary.R")
summary(glht(PM_mv, linfct=cbind(contrMat(rep(1,6), type="Tukey"))), test=adjusted("none"))
contrast_Ssum <- simple_summary(summary(glht(PM_mv, linfct=cbind(contrMat(rep(1,6), type="Tukey"))), test=adjusted("none")))
contrast_Ssum
```

These contrasts can be viewed in a plot
```{r plotContrasts}
par(mar = c(5, 13, 4, 2) + 0.1)
plot(glht(PM_mv, linfct=cbind(contrMat(rep(1,6), type="Tukey"))), yaxt='n')
axis(2, at = seq_along(contrast_Ssum$contrast), labels = rev(contrast_Ssum$contrast), las=2, cex.axis = 0.8)
```

Does disease pressure on the crop affect the optimum spray scenario? 
Above we can see spraying at the first sign of powdery mildew or with in `r min(as.numeric(slimmer_PM_dat[slimmer_PM_dat$spray_management == "Late"| slimmer_PM_dat$spray_management == "Late_plus","fungicide_timing_1"]))` - `r max(as.numeric(slimmer_PM_dat[slimmer_PM_dat$spray_management == "Late"|                   slimmer_PM_dat$spray_management == "Late_plus","fungicide_timing_1"]))` days after first sign is effective at mitigating yield loss.
Lets test this hypothesis by adding it as a moderator and comparing the new model with the original model above.

```{r DisPressComparison}
PM_mv2 <- rma.mv(
   yi = grain_yield.t.ha,
   vi,
   mods = ~ spray_management*D_pres,
   method = "ML",
   random = list(~ spray_management | trial, ~1 | id), 
   struct = "UN",
   control=list(optimizer="optim"),
   data = dat1
)

anova(PM_mv, PM_mv2)
```
We can see that the lower $QE$ indicates that disease pressure moderator helped explain some of the residual heterogeneity in differences in grain yield between treatments. 
However according to the Wald-test the models were not significantly different and we should use the original "Reduced" model.

The P value was close to the 0.05 cut-off so we will present the model output for the interest of the readers.
```{r DissPressMetMod}
summary(PM_mv2)
summary(glht(PM_mv2, linfct=cbind(contrMat(rep(7,12), type="Tukey"))), test=adjusted("none"))
# plot(glht(PM_mv2, linfct=cbind(contrMat(rep(7,12), type="Tukey")))) # I think this plot is too busy to be incorporated
```

@@

Lets present the meta-analysis results for the moderator variables in a forest style plot.
First lets create a table of the results.

@@

```{r metafor_plot}
YieldContrasts <- efficacy %>%
   filter(Treatment != "control") %>% 
   mutate(Treatment = factor(Treatment, levels = rev(
      c(

         "Early",
         "Recommended",
         "Recommended_plus",
         "Late",
         "Late_plus"
      )
   ))) %>%
   ggplot(aes(Treatment, Mean)) +
   geom_hline(
      yintercept = seq(-0.05, 0.3, by = 0.05),
      color = usq_cols("usq charcoal"),
      linetype = 3
   ) +
   geom_hline(yintercept = 0) +
   geom_point(aes(size = 1 / SE), shape = 15) +
   geom_linerange(aes(ymin = CIs_lower, ymax = CI_upper)) +
   coord_flip()

YieldContrasts
```

### Forest plot of trial treatment effect per trial
Lets look at the trial variability in this model with a forest plot
```{r forestP, fig.height= 40, fig.width=17}
# THIS NEEDS TO BE FIXED
source(here("R/change_labels.R"))
source(here("R/forest_rows.R"))

### Sort and arrange data-frame accordingly and produce a index vector to inform the order of each "Trial" on the plot
index1 <- dat1 %>%
   mutate(ID = seq_along(trial_ref)) %>%
   mutate(spray_management = 
             factor(spray_management, 
                    levels = c("Early","Late","Recommended",
                    "Recommended_plus","Late_plus"))) %>%
   arrange(desc(spray_management), location, desc(grain_yield.t.ha)) %>%
   mutate(ID2 = order(ID))%>%
   pull(ID2)

### Use forest rows function to return a list of where to place on the plot: sub-titles, each trial row, and summary prediction for each treatment.
### this list can be used below to reference where to plot each variable
rows1 <- forest_rows(head_row = 186, 
                     ord_var = dat1$spray_management, 
                     index = index1, 
                     gap = 5, 
                     row_offset = -2)

### A data-frame of meta-model coefficients used to produce the summary polygons for each treatment
polys <- data.frame(ref = change_labels(names(coef(PM_mv))),
                    x = c(coef(PM_mv)[1],coef(PM_mv)[1] + coef(PM_mv)[2:6]),
                    ci.lb = c(PM_mv$ci.lb[1],PM_mv$ci.lb[1] +PM_mv$ci.lb[2:6]),
                    ci.ub = c(PM_mv$ci.ub[1],PM_mv$ci.ub[1] +PM_mv$ci.ub[2:6]))[c(2,5,3,4,6,1),]
polys$rows <- rows1$pred_row - 1.5

forest1  <- metafor::forest(PM_mv, cex = 1.15 , width = 5,
                #order = "fit",
                rows = rows1$trows-1.5,
                ilab = cbind(dat1$year, 
                             dat1$row_spacing, dat1$host_genotype, dat1$dose),
                ilab.xpos = c(-2.35,-2.1,-1.8,-1.4)+1,
                slab = NA, 
                xlim = c(-2, 4),
                ylim = c(5,190),
                xlab = "Grain yield mean difference",
                addfit = FALSE,
                efac = c(0.2,0.15,0.2),
                at = seq(-0.5,3, 0.5))

### code which puts lines and annotations on the plot, its a bit long...
source(here("R/add_annotations.R"))

@@

@@

```{r Save_meta_data, eval=FALSE}
write.csv(slimmer_PM_dat, file = "data/GYmeta_data.csv")
save(dat1, efficacy, PM_mv, YieldContrasts, file = here("cache/Meta-analysisData.Rdata"))
```

#### Powdery mildewMBM comparison -- Delete
One aspect we would like to consider is how the results of this meta-analysis compares to how the PowderyMildewMBM estimates the effect of fungicide applications on grain yields.
Below we create a data frame of the mitigation factors for each spray scenario so we can compare them to the outcomes of the meta-analysis.
```{r compare2PowderyMildewMDM}

PMBM <- 
   data.frame(Sprays = c(rep(1,20), rep(2,20)),
              Disease = rep(c(rep("NotPresent", 5),
                          rep("LowerCanopy", 5),
                          rep("MidCanopy", 5),
                          rep("UpperCanopy", 5)),2),
              CropMaturity = rep(c("Vegetative", "Budding", "Flowering",
                                   "EarlyPod", "LatePod"), 8),
              mitigation = c(0.57, 0.855, 0.95,0.95,0.95,
                             0.45995, 0.52919,0.5530,0.390,0.1642,
                             0.291258,0.351051, 0.2454,0.1409,0.4717,
                             0,0.059597,0.093034,0.0682476,0.0215584,
                             0.95,0.95,0.95,0.95,0.95,
                             0.766588,0.6991,0.55042,0.390056,0.1642,
                             0.485429,0.390,0.2454,0.140904,0.0471684,
                             0.0791,0.086068,0.093034,0.0682476,0.0215584))

# what is the mean age of the plant in our data when disease first occurs and we apply the first spray?
source("R/import_data.R")
PM_MB_dat <- import_data()
mean(PM_MB_dat$first_sign_disease - PM_MB_dat$planting_date, na.rm =TRUE) # 46.3 days (which is when the plant is flowering)

#What proportion of the yield was saved?
1 - (coef(PM_mv)[1] /(coef(PM_mv)[1] + coef(PM_mv)[3]))
1- (coef(PM_mv)[1] /(coef(PM_mv)[1] + coef(PM_mv)[4]))

PMBM %>%
   filter(CropMaturity == "Flowering")

```
To to a better comparison I need to know how the mitigation factor is calculated with the estimated yields.
